---
title: "Home runs"
---

```{r}
#| cache: true
#| message: false
library(tidyverse)
library(Lahman)

shortened_seasons <- c(1918, 1919, 1981, 1994, 1995, 2020)

hr_by_season <- Teams |>
  group_by(yearID) |>
  summarise(
    total_hr = sum(HR, na.rm = TRUE),
    total_team_games = sum(G, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    hr_per_game = total_hr / total_team_games,
    shortened = yearID %in% shortened_seasons
  )
```

```{r}
library(tidyverse)

ggplot(hr_by_season, aes(x = yearID, y = hr_per_game, color = shortened)) +
  geom_line(linewidth = 1) +
  scale_color_manual(
    values = c("FALSE" = "steelblue", "TRUE" = "firebrick"),
    labels = c("Regular season", "Shortened season")
  ) +
  labs(
    title = "Home runs per team-game by season",
    x = "Season",
    y = "Home runs per team-game",
    color = "Season type",
    caption = "Source: Lahman Teams dataset; shortened seasons marked in red."
  ) +
  theme_minimal()
```

The suspicious seasons stand out sharply when compared to a normal recent year. In 2019, MLB played 4,858 team-games and produced 6,776 home runs, or about 1.39 home runs per team-game. By contrast, the shorter or disrupted seasons had much lower rates:

```{r}
library(Lahman)

comparison <- Teams |>
  group_by(yearID) |>
  summarise(
    total_hr = sum(HR, na.rm = TRUE),
    total_team_games = sum(G, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(hr_per_game = total_hr / total_team_games) |>
  filter(yearID %in% c(2019, 1981, 1994, 2020)) |>
  arrange(desc(yearID))

comparison |>
  rename(
    Season = yearID,
    `Total home runs` = total_hr,
    `Total team-games` = total_team_games,
    `Home runs per game` = hr_per_game
  ) |>
  knitr::kable(digits = 2, caption = "Comparison of a normal recent season with the suspicious seasons")
```

The drop in home-run rate in 1981, 1994, and 2020 is consistent with those seasons being shortened or affected by disruptions, while 2019 reflects a more typical full season.

This page explores home run trends in Major League Baseball.
